\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-05-02 project-writings/paper.tex
remove the Surprises and Lessons Learned section
Do a bunch of work on the conclusion.

\documentclass{article}
\usepackage{fullpage}
%\usepackage{times}
\usepackage{url}
\usepackage{amsmath,amsthm,amssymb,color}
%\usepackage{multicol}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{courier}
\usepackage{listings}
\usepackage{color}
\definecolor{lightgray}{rgb}{.9,.9,.9}
%\usepackage[all]{hypcap}

@@

\let\x\cdot
\lstset{
  language=Python,
  backgroundcolor=\color{lightgray},
  showstringspaces=false,
  tabsize=4,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\ttfamily,
  captionpos=b
}
\lstdefinelanguage{JavaScript}{
  keywords={if,else,var},
  keywordstyle=\bfseries\ttfamily,
  backgroundcolor=\color{lightgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/}
}

@@

\section{Related Work}
%% XXX: FIXME, TODO
There are a variety of approaches to range analysis. A recent work, ``The Design
and Implementation of a Non-Iterative Range Analysis Algorithm on a
Production Compiler'', by Teixeira and Pereira, outlines a practical implementation of
implementation of an algorithm designed by Su and Wagner
\cite{Su04aclass}. The algorithm due to Su and Wagner is
non-iterative, and works by constructing a constraint graph from the program's control flow graph.

@@

IonMonkey has two intermediate languages. There is a ``Middle
Intermediate Representation'' (MIR) which encompasses higher-level
constructs such as lambdas, objects, and strings. IonMonkey performs a variety of
optimizations including global value numbering and loop invariant code motion,
and then the code is lowered to a ``Lower Intermediate Representation'' (LIR)
used for register allocation and code generation. IonMonkey currently supports
code generation for the x86, x86\_64, and ARM architectures.

@@

\begin{lstlisting}[language=Javascript,
                   caption={Code that requires information about branches to properly analyze},
                   label={fig:branch}, float=ht]
var y;
if (x < 0) {
    y = x + 2000000000;
} else {
    if (x < 1000000000) {
        y = x*2;
    } else {
        y = x - 3000000000;
    }
}
\end{lstlisting}

\begin{lstlisting}[language=Javascript,
                   caption={Code from Figure \ref{fig:branch}, in XSA form},
                   label={fig:xsa}, float=ht]
if (x < 0) {
    x1 = Beta(x, [INT_MIN, -1]);
    y1 = x1 + 2000000000;
} else {
    x2 = Beta(x, [0, INT_MAX]);
    if (x2 < 1000000000) {
        x3 = Beta(x2, [INT_MIN, 999999999]);
        y2 = x3*2;
    } else {
        x4 = Beta(x2, [1000000000, INT_MAX]);
        y3 = x4 - 3000000000;
    }
    y4 = Phi(y2, y3);
}
y = Phi(y1, y4);
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,
                   caption={Pseudocode algorithm for inserting beta nodes},
                   label={lst:add_betas}, float=ht]
def addBetaNodesToFunc(cfg):
  for b in cfg.basic_blocks():
    if b.numPredecessors() == 1 and isBranch(b.pred(0).lastInstruction()):
      test = b.pred(0).lastInstruction()
      # find whether we are the true or false branch
      direction = (b == test.trueTarget())
      if test matches ``x RELOP c'':
          r = computeRange(RELOP, c, direction)
          x_prime = b.insertFront(Beta(x, r))
          replaceDominatedUses(x, x_prime)

@@

\begin{lstlisting}[language=Python,
                   caption={Pseudocode algorithm for propagating range information},
                   label={lst:propagate_ranges}, float=ht]
def propagateRanges(graph):
    worklist = []
    for block in reversePostOrder(graph):
        for insn in block:
            insn.recomputeRange()
            if !insn.isPhi() and !insn.isBeta():
                continue
            worklist.appendAll(insn.uses())

@@

There are three benchmark suites which are generally used to measure JavaScript
JIT compilers: the v8bench suite developed by Google\cite{v8bench}, the SunSpider benchmark
suite developed by Apple\cite{sunspider}, and the Kraken benchmark suite developed by Mozilla\cite{kraken}

We ran each benchmark suite on IonMonkey, both with and without our range
analysis pass, and measured the performance differences. The results are
summarized in Figures \ref{fig:v8bench}, \ref{fig:sunspider}, and
\ref{fig:kraken}. These figures were gathered using the SunSpider test harness,
which gives both a mean and a 95\% confidence interval\cite{sunspider}. Each
test was run 30 times.

>>>>

\documentclass{article}
\usepackage{fullpage}
%\usepackage{times}
\usepackage{url}
\usepackage{amsmath,amsthm,amssymb,color}
%\usepackage{multicol}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{courier}
\usepackage{listings}
%\usepackage[all]{hypcap}

@@

\let\x\cdot
\lstset{
  language=Python,
  showstringspaces=false,
  tabsize=4,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\ttfamily,
  captionpos=b
}

@@

\section{Related Work}
%% XXX: FIXME, TODO
There are a variety of approaches to range analysis. A recent work, ``The Design
and Implementation of a Non-Iterative Range Analysis Algorithm on a
Production Compiler'', by Teixeira and Pereira, outlines a practical implementation of
implementation of an algorithm designed by by Su and Wagner
\cite{Su04aclass}. The algorithm due to Su and Wagner is
non-iterative, and works by constructing a constraint graph from the program's control flow graph.

@@

IonMonkey has two intermediate languages. There is a ``Middle
Intermediate Representation'' (MIR) which encompasses higher-level
constructs such as lambdas, objects, and strings. A variety of
optimizations are performed, and then the code is lowered to a ``Lower
Intermediate Representation'' (LIR) used for register allocation and
code generation.

@@

\begin{figure}[ht]
\begin{center}
\begin{verbatim}
var y;
if (x < 0) {
    y = x + 2000000000;
} else {
    if (x < 1000000000) {
        y = x*2;
    } else {
        y = x - 3000000000;
    }
}
\end{verbatim}
\end{center}
\caption{Code that requires information about branches to properly analyze}
\label{fig:branch}
\end{figure}

\begin{figure}[ht]
\begin{center}
\begin{verbatim}
if (x < 0) {
    x1 = Beta(x, [INT_MIN, -1]);
    y1 = x1 + 2000000000;
} else {
    x2 = Beta(x, [0, INT_MAX]);
    if (x2 < 1000000000) {
        x3 = Beta(x2, [INT_MIN, 999999999]);
        y2 = x3*2;
    } else {
        x4 = Beta(x2, [1000000000, INT_MAX]);
        y3 = x4 - 3000000000;
    }
    y4 = Phi(y2, y3);
}
y = Phi(y1, y4);
\end{verbatim}
\end{center}
\caption{Code from Figure \ref{fig:branch}, in XSA form}
\label{fig:xsa}
\end{figure}

@@

\begin{lstlisting}[language=Python,
                   caption={Pseudocode algorithm for inserting beta nodes},
                   label={lst:add_betas}]
def addBetaNodesToFunc(cfg):
  for b in cfg.basic_blocks():
    if b.numPredecessors() == 1 and isBranch(b.pred(0).lastInstruction()):
      test = b.pred(0).lastInstruction()
      # find whether we are the true or false branch
      direction = (b == test.trueTarget())
      if test matches ``x RELOP c'':
          r = computeRange(RELOP, c, direction)
          x_prime = b.insertFront(Beta(x, r))
          replaceDominatedUses(x, x_prime)

@@

\begin{lstlisting}[language=Python,
                   caption={Pseudocode algorithm for propagating range information},
                   label={lst:propagate_ranges}]
def propagateRanges(graph):
    worklist = []
    for block in reversePostOrder(graph):
        for insn in block:
            insn.recomputeRange()
            if !insn.isPhi() and !insn.isBeta():
                continue
            worklist.appendAll(insn.uses())

@@

There are three benchmark suites which are generally used to measure JavaScript
JIT compilers: the v8bench suite developed by Google, the SunSpider benchmark
suite developed by Apple, and the Kraken benchmark suite developed by Mozilla.

We ran each benchmark suite on IonMonkey, both with and without our range
analysis pass, and measured the performance differences. The results are
summarized in Figures \ref{fig:v8bench}, \ref{fig:sunspider}, and \ref{fig:kraken}.
